---
title: "Data list"
output:
  rmarkdown::html_document:
    toc: true
    toc_float: true
pkgdown:
  as_is: true
---

## Raw data

This page is a listing of the raw data files in the {dpjr} package. The files can be accessed using the appropriate "read" function, with the package function `dpjr_data()` providing the path. For example, 

```
df_mpg <- read.csv(dpjr_data("mpg.csv"))
```

The full list of files is as follows:

```{r data_list}
dpjr::dpjr_data()
```


## Licenses

All of the data in this package is covered under an open license. For information about the specific license for each data set in this package, please refer to [the vignette "Data licenses"](data_licenses.html). 



## Files

### authors_fwf.txt; authors2_fwf.txt

A pair of fixed-width files containing information about ten authors from the United States of America: 

* "name"

* "state of birth", the two-letter abbreviation for the US state in which the author was born

* "unique_id", a randomly generated personal ID (similar to a national individual identification number used in different countries).


The variables are as follows:

Variable           Width   Start position   End position
--------           -----   --------------   ------------
name               20      1                20
state_of_birth     10      21               30
unique_id          12      31               42



### cr25

The files within the subdirectory `cr25` are as follows:

```{r data_list_cr25}
list.files(dpjr::dpjr_data("cr25"))
```

For information about these files see [the vignette "Data generation—CR25 orders"](data_generate_cr25_order.html).


### gapminder.csv

A csv version of the dataset in the [{gapminder}](https://CRAN.R-project.org/package=gapminder) package


### JCUSH.txt

Statistics Canada has made available an anonymized Public-Use Microdata File (PUMF) of the Joint Canada/United States Survey of Health, a telephone survey conducted in late 2002 and early 2003. There were 8,688 respondents to the survey, 3,505 Canadians and 5,183 Americans. 

The PUMF is a fixed-width file named "JCUSH.txt". Each line is 552 columns in length.


The webpage for the survey, including the PUMF file, data dictionary, and methodological notes, is here:
https://www150.statcan.gc.ca/n1/pub/82m0022x/2003001/4069119-eng.htm



**License:**

[Statistics Canada Open License](https://www.statcan.gc.ca/en/reference/licence)

Source: Statistics Canada, Joint Canada/United States Survey of Health 2002-03, 2004. Reproduced and distributed on an "as is" basis with the permission of Statistics Canada.



### mpg.csv

A csv version of the famous mpg dataset, included in the [{ggplot2} package](https://ggplot2.tidyverse.org/reference/mpg.html)


### penguins_fwf.txt

A fixed-width version of the [{palmerpenguins} dataset](https://allisonhorst.github.io/palmerpenguins/)

There are 8 different variables, described in the table below:

Variable           Width   Start position   End position
--------           -----   --------------   ------------
species            9       1                9
island             9       10               18
bill_length_mm     4       19               22
bill_depth_mm      4       23               26
flipper_length_mm  3       27               29
body_mass_g        4       30               33
sex                6       34               39
year               4       40               43




The fixed-width file has been created to minimize white space. The first four and last two rows of the data look like this:


```{r penguin_readLines_head}
readLines(dpjr::dpjr_data("penguins_fwf.txt"), n = 4)
```

```{r penguin_readLines_tail}
tail(readLines(dpjr::dpjr_data("penguins_fwf.txt")), 2)
```

Note that the first row is _not_ the variable names. This is common in fixed-width files.


**License:**

Data are available by CC-0 license in accordance with the Palmer Station LTER Data Policy and the LTER Data Access Policy for Type I data.



